---
title: "pubh7462_hw2"
author: "Xiao Ma"
output: github_document
---
# 3.1 Data Exploration & Cleaning
```{r message=FALSE, warning=FALSE}
library(DataExplorer)
library(tidyverse)
library(scales)
library(ggplot2)
library(gt)
library(ggrepel)
library(paletteer)
BRFSS <- read.csv("./data/brfss_smart_2010.csv", stringsAsFactors=TRUE)
#names(BRFSS)
brfss2 <- BRFSS %>%
  separate(col = Locationdesc, into = c("State","2","County"), sep = " ") %>%
  mutate(Prop_response=Data_value) %>%
  filter(Topic=="Overall Health") %>%
  select(Year,State,County,Response,Sample_Size,Prop_response)
brfss2$Year<-as.factor(brfss2$Year)
brfss2$State<-as.factor(brfss2$State)
brfss2$County<-as.factor(brfss2$County)
```

# 3.2 Data Description

- There are 10625 observations and 6 variables after data cleaning.
- Each case is a unique contain the Year, State, County, Response, Sample size, Proportion of response.
- Year, State, County, Response are quite intuitive and straightforward, Response describes the content of response, Sample size describes the observed locations, Prop_response states the Proportion of response when respondents filling in the questionnaire.
- Each one observed locations have more than one response.

# 3.3 Do Data Science
# 3.3.1
```{r message=FALSE, warning=FALSE}
brfss2 %>%
  filter(Year==2004) %>%
  distinct(State, County)%>%
  group_by(State) %>%
  summarise(Locations=n()) %>%
  filter(Locations==6) %>%
  gt() %>%
  tab_header(title = "States observed at 6 locations") %>%
  tab_source_note(source_note = "Data source: Behavioral Risk Factors Surveillance System")
```

- CO, NM, SC, TX, UT, VT are states observed at 6 locations.

# 3.3.2
```{r message=FALSE, warning=FALSE}
brfss3 <- brfss2 %>%
  distinct(State, County,Year) %>%
  group_by(Year,State) %>%
  summarise(Oberved_location=n()) %>%
  ungroup(Year,State) %>%
  mutate(State=fct_reorder(State,Oberved_location,mean,.desc = TRUE)) 

brfss3 %>%
  ggplot(aes(x=Year,y=Oberved_location,group=State,color=State)) +  
  geom_point(color = "grey", size = 0.5) +
  geom_line(size = 0.3, alpha = 0.8) +
  geom_line(data =brfss3[brfss3$State=="NJ",],
            aes(x=Year,y=Oberved_location), 
            color = "dodgerblue4", size = 0.6) + 
  geom_label(label="New Jersey", x=2004,y=24.2,label.padding = unit(0.55, "lines"), 
    label.size = 0.35,
    color = "dodgerblue4",
    fill="white")+
  xlab("")+
  ylab("Observed number")+
  theme_light()+
  labs(title = "Number of observed locations in each state",
       subtitle = "From 2002 to 2010",
        caption = "Data from the Behavioral Risk Factors Surveillance System (BRFSS)")
```

- There are some fluctuations of each states during years, but the line do not have a clear increasing or decreasing pattern.
- It's noteworthy that Florida had distinct change during the years. 

# 3.3.3
```{r message=FALSE, warning=FALSE}
brfss4 <- brfss2 %>%
  filter(State=="MN",Year %in% c(2002,2006,2010),Response %in% c("Excellent","Good","Poor")) %>%
  group_by(Year,Response) %>%
  summarise(across(c(Sample_Size,Prop_response),  list(mean = mean, sd = sd), na.rm=FALSE, .names = "{.col}.{.fn}"))

brfss4 %>%
  gt() %>%
  fmt_number(columns = vars(Sample_Size.mean,Sample_Size.sd,Prop_response.mean,Prop_response.sd), 
             decimals = 2) %>%
  fmt_percent(columns = vars(Prop_response.mean), 
              decimals = 2, 
              scale_values = FALSE)%>%
  tab_header(title = "Mean and standard deviation of responses",subtitle = "Years of 2002, 2006, 2010, Responses: sample size and proportion") %>%
  tab_source_note(source_note = "Data source: Behavioral Risk Factors Surveillance System")
```

- For mean of proportion of response and sample size, it is obvious that group of excellent and good have larger mean than poor group. The plot have an increasing trend.
- Sd of Prop_response of group of excellent is larger than that of good and poor group.
- Sd of Sample size of group of excellent and good are really close and is larger than that of poor group.

# 3.3.4
```{r message=FALSE, warning=FALSE}
brfss5<-brfss4 %>%
  pivot_longer(cols = c(Sample_Size.mean,Sample_Size.sd,Prop_response.mean,Prop_response.sd),names_to = "Statistics",values_to = "Value")

brfss5 %>%
  ggplot(aes(x=Year,y=Value,group=Response,color=Response)) +
  geom_point() +
  geom_line()+
  facet_wrap(~Statistics, scales = "free")+
  xlab("")+
  ylab("Observed number")+
  theme_light()+
  labs(title = "Mean and standard deviation of responses",subtitle = "Years of 2002, 2006, 2010",
        caption = "Data from the Behavioral Risk Factors Surveillance System (BRFSS)")
```

